NumPy---Array Programming in Python

### 2020-01-20 body.tex
Rethinking Fig. 2 (duck array / interoperability)

\begin{figure*}
  \centering
  \includegraphics[width=0.6\textwidth]{static/sketches/duck-arrays}
  \caption{\textbf{Proliferation of Arrays.} The NumPy array is a classic and simple in-memory representation of array data. Since its inception, numerous other array representations have been developed, including distributed arrays---which are stored on multiple computers, GPU arrays---which utilize graphics processing units for fast computation, computation graphs---which postpone and combine calculations before executing them efficiently, and sparse arrays---which typically contain few non-zero values, and store only those in memory.}
  \label{fig:array-proliferation}
\end{figure*}

Recent years have seen a massive explosion in data science, machine learning, and artificial intelligence.  NumPy and its API, underlying almost every tool in the computational toolchain, has become ubiquitous.  Still, NumPy itself could never hope to satisfy every single specialized need of the community; it is limited in its ability to work with very large datasets, datasets split across multiple systems, and computations that require specialized hardware.

Furthermore, pursuing solutions to these challenges is out of scope for
NumPy development. The community's efforts to fill the resultant gaps therefore
led to a proliferation of arrays (Fig.~\ref{fig:array-proliferation}):
distributed arrays, GPU arrays, delayed execution arrays, sparse arrays,
foreign language arrays, and more.

@@

To facilitate \emph{interoperability}, the NumPy team has designed
``protocols'' (or contracts of operation), that allow for these arrays to be
passed to NumPy functions.
NumPy, in turn, dispatches operations to the originating library, as required.
% FIXME: expand to 3 or more sentences discussing current efforts and
% issues with those, how did adoption go, what is being done next
%\url{https://numpy.org/neps/nep-0037-array-module.html}
Versions of these protocols have been successfully deployed.
% Ralf: Could you provide some proof of effectiveness, e.g., "This has made it
% possible to create distributed GPU arrays, enabling .... [find a reference to
% work by Peter Entschev]".

@@

% S: I think two issues are conflated here: the simple underlying memory model, and the API.  The API has developed organically, while the underlying memory model is a design decision that shapes much of how the rest of the library was built.

% S:
%
% Thinking about enabling factors, I'd characterize them into three categories:
%
% - Practical
% - Philosophical
% - Social
%
% E.g., practical: reason why we use Python (learn one language for everything); students don't have money, so want to avoid impracticalities of license dongles/servers, etc.
%
% Philosophical: science should be open, transparent; our software should be controlled by scientists, not designers that we don't have access to
%
% Social: joy of building these things together, friendly welcome into the community for many of us---appreciation of our work and hours
%

>>>>

Recent years have seen a massive explosion in data science, machine learning,
and artificial intelligence.  NumPy and its API, underlying almost every tool
in the computational toolchain, has become ubiquitous.  Still, NumPy itself
could never hope to satisfy every single specialized need of the community; it
is limited in its ability to work with very large datasets, datasets split
across multiple systems, and computations that require specialized hardware.
Furthermore, pursuing solutions to these challenges is out of scope for NumPy
development.

The community's efforts to fill the resultant gaps therefore led to a
proliferation of arrays including distributed arrays---which are stored on
multiple computers, GPU arrays---which utilize graphics processing units for
fast computation, computation graphs---which postpone and combine calculations
before executing them efficiently, and sparse arrays---which typically contain
few non-zero values, and store only those in memory.
%distributed arrays, GPU arrays, delayed execution arrays, sparse arrays,
%foreign language arrays, and more.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{static/sketches/duck-arrays}
  \caption{\textbf{Interoperability.} \fixme{This figure needs work.}}  \label{fig:duck-arrays}
\end{figure*}

To facilitate \emph{interoperability}, the NumPy team has designed
``protocols'' (or contracts of operation), that allow for these arrays to be
passed to NumPy functions (Fig.~\ref{fig:duck-arrays}).
NumPy, in turn, dispatches operations to the originating library, as required.
% FIXME: expand to 3 or more sentences discussing current efforts and
% issues with those, how did adoption go, what is being done next
%\url{https://numpy.org/neps/nep-0037-array-module.html}
Versions of these protocols have been successfully deployed.
% Ralf: Could you provide some proof of effectiveness, e.g., "This has made it
% possible to create distributed GPU arrays, enabling .... [find a reference to
% work by Peter Entschev]".

@@

% S:
%
% Thinking about enabling factors, I'd characterize them into three categories:
%
% - Practical
% - Philosophical
% - Social
%
% E.g., practical: reason why we use Python (learn one language for
% everything); students don't have money, so want to avoid impracticalities of
% license dongles/servers, etc.
%
% Philosophical: science should be open, transparent; our software should be
% controlled by scientists, not designers that we don't have access to
%
% Social: joy of building these things together, friendly welcome into the
% community for many of us---appreciation of our work and hours
